Community Detection Methods for the Generalized Random Dot Product Graph Model

### 2021-05-24 proposal/slides.Rmd
updated slides

```{r, fig.height = 4, fig.width = 4, out.width = '50%'}
P.eigen = eigen(P)
X <- P.eigen$vectors[, 1:2] %*% diag(P.eigen$values[1:2] ** .5)

@@

$A$ is sampled from a PABM if 
$A \sim BernoulliGraph(P)$ such that:

1. Let $P^{(kl)}$ denote the $n_k \times n_l$ matrix of edge probabilities 
between communities $k$ and $l$. 
2. Organize popularity parameters as vectors 
$\lambda^{(kl)} \in \mathbb{R}^{n_k}$ 
such that $\lambda^{(kl)}_i$ is the popularity parameter 
of the $i$^th^ vertex of community $k$ towards community $l$. 
3. Each block can be decomposed as 
$P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$.

@@

**Corollary**: $X$ is block diagonal by community and 
$U$ is orthogonal $\implies$ 
each community corresponds to a subspace in $\mathbb{R}^{K^2}$.

@@

If $X$ obeys the *Subspace Detection Property*, then $B$ is sparse such that 
$B_{ij} = 0$ if $i$ and $j$ belong to different communities and $\|c_i\| > 0$.

@@

@@

$$\arg\min_{c_i} \|c_i\|_1 \text{ subject to } A_{\cdot, i} = A c_i
\text{ and } c_i^{(i)} = 0$$

@@

$$\arg\min_{c_i} \|c_i\|_1 \text{ subject to } 
\hat{x}_i = \hat{X} c_i \text{ and } c_i^{(i)} = 0$$
$$A \approx \hat{X} I_{p, q} \hat{X}^\top$$

@@

1. Each point's distance to its subspace is sufficiently small.
2. The cosine of the angles between pairs of subspaces is sufficiently small.
3. The points corresponding to each subspace cover a sufficient amount of 
that subspace.

>>>>

```{r, fig.height = 4, fig.width = 5, out.width = '50%'}
P.eigen = eigen(P)
X <- P.eigen$vectors[, 1:2] %*% diag(P.eigen$values[1:2] ** .5)

@@

$A$ is sampled from a PABM if $P$ can be described as:

1. Let each $P^{(kl)}$ denote the $n_k \times n_l$ matrix of edge probabilities 
between communities $k$ and $l$. 
2. Organize popularity parameters as vectors 
$\lambda^{(kl)} \in \mathbb{R}^{n_k}$ 
such that $\lambda^{(kl)}_i = \lambda_{k_i l}$ is the popularity parameter 
of the $i$^th^ vertex of community $k$ towards community $l$. 
3. Each block can be decomposed as 
$P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$.

@@

**Corollary**: Since $X$ is block diagonal by community and $U$ is orthogonal,
each community corresponds to a subspace in $\mathbb{R}^{K^2}$.

@@

@@

If $X$ obeys the *Subspace Detection Property*, then $B$ is sparse such that 
$B_{ij} = 0$ if $x_i$ and $x_j$ belong to different subspaces

```{r, fig.height = 5, fig.width = 5, out.width = '75%'}
set.seed(314159)

x1 <- runif(n, 1 / sqrt(50), 1)
y1 <- rep(0, n)
x2 <- rep(0, n)
y2 <- x1
z <- c(rep(1, n), rep(2, n))
X <- cbind(c(x1, x2), c(y1, y2))
X <- X + cbind(runif(n * 2, 0, .025),
               runif(n * 2, 0, .025))
X <- X %*% rbind(c(cos(pi / 9), sin(pi / 9)),
                 c(-sin(pi/ 9), cos(pi / 9)))
plot(X, col = z * 2, asp = 1, xlab = NA, ylab = NA)
```

```{r, fig.height = 5, fig.width = 5, out.width = '75%'}
Y <- t(X)
Y <- sweep(Y, 2, apply(Y, 2, function(y) sqrt(sum(y ** 2))), `/`)
Y[is.nan(Y)] <- 0
lambda <- .05
N <- ncol(Y)
B <- plyr::aaply(seq(N), 1, function(i) {
  y <- Y[, i]
  X <- Y[, -i]
  betahat <- glmnet::glmnet(X, y, lambda = lambda, intercept = FALSE) %>% 
    coef() %>% 
    as.numeric()
  if (i != N) {
    betahat <- c(betahat[seq(i)], 0, betahat[seq(i + 1, N)])
  } else {
    betahat <- c(betahat, 0)
  }
  betahat <- betahat[-1]
  return(betahat)
}, .parallel = TRUE) %>% 
  abs()
B <- sweep(B, 2, apply(B, 2, max), `/`)
B[is.nan(B)] <- 0
W <- B + t(B)

qgraph::qgraph(sign(W), layout = 'spring', groups = factor(z))
```

@@

$$c_i = \arg\min_{c} \|c\|_1 \text{ subject to } A_{\cdot, i} = A c
\text{ and } c^{(i)} = 0$$

@@

$$c_i = \arg\min_{c} \|c\|_1 \text{ subject to } 
\hat{x}_i = \hat{X} c \text{ and } c^{(i)} = 0$$
$$A \approx \hat{X} I_{p, q} \hat{X}^\top$$

@@

1. Each point's distance to its subspace is sufficiently small.
2. The points corresponding to each subspace cover a sufficient amount of 
that subspace.
3. The cosine of the angles between pairs of subspaces is sufficiently small.
